Erroneous use of Venn diagrams leads to artifacts in transcriptomic analyses

### 2021-11-19 mk2/manuscript.rmd
small fixes

---
title: "Erroneous use of Venn diagram leads to artifacts in transcriptomic analyses"
author: "January Weiner^1,†^, Benedikt Obermayer^1^ and Dieter Beule^1^"
date: "`r Sys.Date()`"
outputoff:
  html_document
output:
  word_document:
    reference_docx: templates/template.docx
  pdf_document:
    latex_engine: xelatex
bibliography: manuscript.bib
link-citations: true
XXcsl: templates/nature-medicine.csl
---

@@

Venn diagrams are frequently used to illustrate the results of
differential expression analysis and further
downstream functional analysis. Here we show that both, the use
of Venn diagrams to find genes which are thought to be specific for a
certain comparison, as well as gene set enrichment analysis applied to
such subsets is a fallacy. Moreover, the combination of incorrect statistical reasoning
and gene set enrichment analysis may result in particularly misleading artifacts.

>>>>

---
title: "Erroneous use of Venn diagrams leads to artifacts in transcriptomic analyses"
author: "January Weiner^1,†^, Benedikt Obermayer^1^ and Dieter Beule^1^"
date: "`r Sys.Date()`"
outputoff:
  html_document
output:
  word_document:
    reference_docx: templates/template.docx
outputoff2: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
bibliography: manuscript.bib
link-citations: true
XXcsl: templates/nature-medicine.csl
---

@@

Venn diagrams are frequently used to illustrate the results of
differential expression analysis and further
downstream functional analysis. Here we show that both, the use
of Venn diagrams to find genes which are thought to be specific for a
certain comparison, as well as gene set enrichment analysis applied to such
subsets is a fallacy. Since genes which show a significant change in one
condition but not in another are likely to be false negatives in the latter
case, these genes considered to be specific for one condition are likely
characteristic for both of the compared groups, and enriched in functions
related to the research hypothesis. Thus, the combination of incorrect
statistical reasoning and gene set enrichment analysis may result in
particularly misleading artifacts.
